---
title: "Aides Drug"
author: "Jose Trejo"
date: "put the date here"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides (triUmpsal) is under test in clinical trials. Analyze the effectiveness of this drug across the board. Consider population subgroups by gender, drug use, and sexual orientation. Finish this!

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
All variables in this study are categorical as a result we will use bar chart for grahpical display a cross table to summarize numerical results and a chi square and or fisher test for inference. 
## Question 1: is the drug effective in the general population?

The first question is whether or not the vaccine was effective in the whole population. 

$H_O$ The drug has no effect on whether or not one gets AIDS.

$H_A$ The drug does have an effect on whether or not one contracts AIDS.

#Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment, tyoe ='percent', data=triUmp)
```

This barchart shows the vaccine may have an effect on whether one contracts AIDS.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment, data=triUmp)
rowPerc(table1)
colPerc(table1)
```

row percentages on top and table of columns percentsges on bottom.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

For chi square test the p value was 0.0000002 which is less than 0.5, which is evidence to reject the null hypothesis. 


## Question 2: Is the drug effective in the LGBTQ community. 

$H_O$ : The drug has no effect on the LGBTQ community contracting AIDS. 

$H_A$ : The drug has an effect on the LGBTQ community contracting AIDS. 

#Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
LGBTQdata = filter(triUmp, LGBTQ != "straight")
head(LGBTQdata)
tail(LGBTQdata)

barchartGC(~infected + treatment, data = LGBTQdata, type = 'percent')
```

The barchart shows that the drug does not have much effect in the LGBTQ community.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~infected + treatment, data = LGBTQdata)
rowPerc(table2)
colPerc(table2)
```

The table shows the drug was not too effective. 
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

chi square test the p value was .0000002 which is less than .05 so we reject the null hypothesis. 

## Question 3: Is the drug effective in the Male population. 

$H_O$ The drug has no effect on whether or not males contract AIDS.

$H_A$ The drug does have an effect on whether or not malesd contract AIDS.

#Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
MalesData = filter(triUmp, gender =="M")
barchartGC(~infected + treatment, type = 'percent',data = MalesData)
```

Barchart shows drug may have an effect on the male population.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table3 <- xtabs(~infected + treatment, data=MalesData)
rowPerc(table3)
colPerc(table3)
```

table shows row percentages on top half and column percentages on other half. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

Chi square shows p value as .0000002 which is less than 5 so we reject the null hypothesis. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
Summary of Results: 
Question 1: The drug is effective in preventing the contraction of AIDS and recommend the drug be used in the gen. population. 

Question 2: The drug does not effect whether one is in the LGBTQ community or not. 

Question 3:If you take the drug as a male you are less likely to contract AIDS.